---
title: "Data dictionary: Patterns of language switching and bilingual childrens word learning: An experiment across two communities"
author: "Rachel K.Y. Tsui, Jessica E. Kosie, Laia Fibla, Casey Lew-Williams, & Krista Byers-Heinlein"
date: '`r format(Sys.time(), "%a %b %d %X %Y")`'
output:
  html_document:
    toc: true
    toc_depth: 4
    toc_float: true
    code_folding: 'hide'
    self_contained: true
  pdf_document:
    toc: yes
    toc_depth: 4
    latex_engine: xelatex
---

```{r setup, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(
  warning = TRUE, # show warnings during codebook generation
  message = TRUE, # show messages during codebook generation
  error = TRUE, # do not interrupt codebook generation in case of errors,
                # usually better for debugging
  echo = TRUE  # show R code
)
ggplot2::theme_set(ggplot2::theme_bw())
pander::panderOptions("table.split.table", Inf)
```

```{r library, message=FALSE, warning=FALSE}
# load libraries
library(codebook)
library(here)
library(dplyr)
library(tidyverse)
library(future)
library(labelled)
```

This is a data dictionary for the data set used in the paper "Patterns of language switching and bilingual childrens word learning: An experiment across two communities".

```{r message=FALSE, warning=FALSE}
#load dataset
data_clean <- read.csv(here("analysis/data_clean.csv"))

```

# **Variable overview**
```{r}
codebook_items(data_clean)
```

# **Subject information**
* *subject_id*: Unique participant ID
* *lang_comm*: The bilingual language community to which the participant belongs; 2 levels: French-English & Spanish-English
* *gender*: 2 levels: Female & Male
* *age_in_years*: Participant's age in years
* *age_in_months*: Participant's age in months
* *parent_edu_level*: Highest education level of participant's parent
* *global_exposure_eng*: Parent-reported global exposure to English (out of 100%)
* *global_exposure_fr*: Parent-reported global exposure to French (out of 100%)
* *global_exposure_sp*: Parent-reported global exposure to Spanish (out of 100%)
* *global_exposure_other*: Parent-reported global exposure to another language (out of 100%)
* *comprehension_proficiency_eng*: Parent-reported English comprehension proficiency score (out of 10)
* *comprehension_proficiency_fr*: Parent-reported French comprehension proficiency score (out of 10)
* *comprehension_proficiency_sp*: Parent-reported Spanish comprehension proficiency score (out of 10)

### Number of participants
This dataset included `r data_clean %>% filter(lang_comm == "French-English") %>% distinct(subject_id) %>% nrow()` French-English bilingual children (`r data_clean %>% filter(gender == "Female" & lang_comm == "French-English") %>% distinct(subject_id) %>% nrow()` girls) and `r data_clean %>% filter(lang_comm == "Spanish-English") %>% distinct(subject_id) %>% nrow()` Spanish-English bilingual children (`r data_clean %>% filter(gender == "Female" & lang_comm == "Spanish-English") %>% distinct(subject_id) %>% nrow()`) who participated in our experiment. 

### Mean age (in months) of participant
```{r message=FALSE, warning=FALSE}
data_clean %>%
  distinct(subject_id, .keep_all = T) %>%
  group_by(lang_comm) %>%
  summarize(mean_age_in_months = mean(age_in_months, na.rm = T),
            sd = sd(age_in_months, na.rm = T),
            min = min(age_in_months, na.rm = T),
            max = max(age_in_months, na.rm = T))
```

### Distribution of parents' education level
```{r message=FALSE, warning=FALSE}
data_clean %>%
  select(subject_id, lang_comm, parent_edu_level) %>%
  distinct(subject_id, .keep_all = T) %>%
  group_by(lang_comm) %>%
  mutate(total_n = length(subject_id)) %>%
  select(lang_comm, parent_edu_level, total_n) %>%
  group_by(lang_comm, parent_edu_level) %>%
  mutate(n = length(parent_edu_level),
         percent = round(n/total_n*100)) %>%
  distinct(parent_edu_level, .keep_all = T)

```

### Mean proficiency
```{r message=FALSE, warning=FALSE}
individual_exposure <- data_clean %>%
  select(subject_id, lang_comm, matches("exposure")) %>%
  distinct(subject_id, .keep_all = T) %>%
  mutate(global_exposure_fr_sp = case_when(lang_comm == "French-English" ~ as.numeric(global_exposure_fr),
                                           lang_comm == "Spanish-English" ~ as.numeric(global_exposure_sp),
                                           TRUE ~ NA_real_)) %>%
  select(subject_id, lang_comm, global_exposure_eng, global_exposure_fr_sp) %>%
  pivot_longer(-c(subject_id, lang_comm), names_to = "language", values_to = "exposure") %>%
  mutate(language = recode(language, "global_exposure_eng" = "English", "global_exposure_fr_sp" = "French/Spanish"))

data_clean %>%
  select(subject_id, lang_comm, matches("exposure")) %>%
  distinct(subject_id, .keep_all = T) %>%
  mutate(global_exposure_fr_sp = case_when(lang_comm == "French-English" ~ as.numeric(global_exposure_fr),
                                           lang_comm == "Spanish-English" ~ as.numeric(global_exposure_sp),
                                           TRUE ~ NA_real_)) %>%
  select(subject_id, lang_comm, global_exposure_eng, global_exposure_fr_sp) %>%
  pivot_longer(-c(subject_id, lang_comm), names_to = "language", values_to = "exposure") %>%
  mutate(language = recode(language, "global_exposure_eng" = "English", "global_exposure_fr_sp" = "French/Spanish")) %>%
  group_by(lang_comm, language) %>%
  summarize(mean_exposure = mean(exposure, na.rm = T),
            sd_exposure = sd(exposure, na.rm = T),
            min_exposure = min(exposure, na.rm = T),
            max_exposure = max(exposure, na.rm = T),
            se = sd_exposure / sqrt(n()),
            CI_lower = mean_exposure - (1.96 * se),
            CI_upper = mean_exposure + (1.96 * se)) %>%
  rename(exposure = mean_exposure) %>%
  ggplot(aes(x = lang_comm, y = exposure, fill = language, color = language)) +
  geom_bar(stat="identity", position=position_dodge(width=0.9), alpha = .25) +
  geom_errorbar(aes(ymin = CI_lower, ymax = CI_upper), width=.2, position = position_dodge(.9), color = "#4d4d4d", alpha = 0.75) +
  geom_dotplot(data = individual_exposure, aes(y = exposure, x = lang_comm, fill = language, color = language), 
               binaxis = 'y', stackdir = 'center', position = position_dodge(0.9), stackratio = 1, dotsize = 0.5, alpha = 0.8) +
  coord_flip()
  
```

# **Exclusion criteria**
* *missing_LEQ*: 1 vs. 0; 1 = participant missing language comprehension proficiency data

```{r message=FALSE, warning=FALSE}
data_clean %>%
  filter(missing_LEQ == 1) %>%
  distinct(subject_id) %>%
  nrow()
```

* *exclude_preterm*: 1 vs. 0; 1 = preterm

```{r message=FALSE, warning=FALSE}
data_clean %>%
  filter(exclude_preterm == 1) %>%
  distinct(subject_id) %>%
  nrow()
```

* *exclude_language_problem*: 1 vs. 0; 1 = parent-reported language or speech development problem

```{r message=FALSE, warning=FALSE}
data_clean %>%
  filter(exclude_language_problem == 1) %>%
  distinct(subject_id) %>%
  nrow()
```

* *exclude_incomplete*: 1 vs. 0; 1 = incomplete data

```{r message=FALSE, warning=FALSE}
data_clean %>%
  filter(exclude_incomplete == 1) %>%
  distinct(subject_id) %>%
  nrow()
```

* *exclude_parent*: 1 vs. 0; 1 = parent-reported technical problems during the experiment

```{r message=FALSE, warning=FALSE}
data_clean %>%
  filter(exclude_parent == 1) %>%
  distinct(subject_id) %>%
  nrow()
```

Note that those paritcipants being flagged as 1 were excluded from the final analysis.

# **Experimental variables**

* *test_order*: Which of the 8 orders did the participant complete
  + English first, immediate-translation first (then French / then Spanish)
  + English first, one-language-at-a-time first (then French / then Spanish)
  + French first, immediate-translation first
  + French first, one-language-at-a-time first
  + Spanish first, immediate-translation first
  + Spanish first, one-language-at-a-time first
```{r message=FALSE, warning=FALSE}
data_clean %>%
  distinct(test_order)
```

* *block*: 3 levels; familiar vs. block1 vs. block2
```{r message=FALSE, warning=FALSE}
data_clean %>%
  distinct(block)
```

* *test_part*: 3 levels; fam_test vs. nov_learn vs. nov_test
  + fam_test refers to the familiar word test trials
  + nov_learn refers to the novel word learning trials
  + nov_test refers to the novel word test trials
```{r message=FALSE, warning=FALSE}
data_clean %>%
  distinct(test_part)
```

* *language*: Which language the trial was in
```{r message=FALSE, warning=FALSE}
data_clean %>%
  group_by(test_part) %>%
  distinct(language) %>%
  arrange(test_part)
```

* *condition*: Which condition the trial was in
  + 3 levels; familiar vs. immediate_translation vs. one_language_at_a_time
```{r message=FALSE, warning=FALSE}
data_clean %>%
  distinct(condition)
```

* *trial_index*: Trial number
* *stimulus*: Which stimulus was played in the trial
* *button_correct*: The correct response for the trial
* *button_selected*: Which response did the participant selected
* *rt*: The response time for the trial
* *accuracy*: If button_correct = button_selected (i.e., the participant selected the correct response), then a score of 1 was given to the trial; if not, a score of 0 was given. If no response was made, then NA was assigned.

### Number of participants per test order 
```{r message=FALSE, warning=FALSE}
data_clean %>%
  select(subject_id, lang_comm, test_order) %>%
  distinct(subject_id, .keep_all = T) %>%
  group_by(lang_comm) %>%
  count(test_order) %>%
  ggplot(aes(x = lang_comm, y = n, fill = test_order)) +
  geom_bar(stat="identity", position=position_dodge(width=0.9)) +
  scale_y_continuous(limits = c(0, 10), breaks = seq(0, 10, by = 1))
```

### Number of trials per participants
```{r message=FALSE, warning=FALSE, fig.height=50, fig.width=20}
level_order <- c("familiar", "block1", "block2")

data_clean %>%
  group_by(subject_id, block) %>%
  filter(test_part == "fam_test" | test_part == "nov_test") %>%
  count() %>%
  ggplot(aes(x = subject_id, y = n, fill = factor(block, level = level_order))) +
  geom_bar(stat="identity", position=position_dodge(width=0.9)) +
  scale_y_continuous(limits = c(0, 12), breaks = seq(0, 12, by = 1)) +
  facet_grid(factor(block, level = level_order) ~ .) +
  coord_flip() +
  theme(text = element_text(size=12),
        axis.text = element_text(size = 12, color = "black"),
        legend.text = element_text(size = 12),
        legend.position="bottom")
```